---
title: "Unit 8 Presentation"
author: "Michael Hedge and Morgan Nelson"
date: "February 17, 2020"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Force installing packages on other people's systems is considered rude.
# If you do not have the required libraries, you may uncomment the following
# section to auto-install them if you so desire.

# if(!require(tidyverse)) {install.packages("tidyverse")}
# if(!require(usmap)) {install.packages("usmap")}
# if(!require(DMwR)) {install.packages("DMwR")}
# if(!require(GGally)) {install.packages("GGally")}
# if(!require(extrafont)) {install.packages("extrafont")}
# if(!require(openintro)) {install.packages("openintro")} # abbr2state
# if(!require(png)) {install.packages("png")} 
# if(!require(jpeg)) {install.packages("jpeg")}
# if(!require(ggpubr)) {install.packages("ggpubr")}


library(tidyverse)
library(usmap)
library(DMwR)
library(GGally)
library(extrafont)
library(openintro)
library(png)
library(jpeg)
library(stringi)
library(ggpubr)
library(class)

loadfonts(quiet = TRUE)
```
Greetings Mr. CEO and Mr. CFO, other distinguished guests.

Given the Beer and Brewery data, our research intends to answer several compelling questions for the business:
1. How many breweries are present in each state?
```{r import beer data}
beers <- read.csv("Beers.csv")
breweries <- read.csv("Breweries.csv")

# Question 2
# Merge beer data with the breweries data. 
# Print the first 6 observations and the last six observations to check the merged file.

# Merge the two data frames into one
data <- merge(beers,breweries,by.x = "Brewery_id", by.y = "Brew_ID")
data <- data %>% 
  rename(Beer = Name.x) %>%       # clean up munged names from merge
  rename(Brewery = Name.y) %>%    # clean up munged names from merge
  rename(Brewery_ID = Brewery_id) # Make anything with _ID the same

head(data, n = 6)                 # First six from merged set
tail(data,n = 6)                  # Last six from merged set

```

```{r Breweries per State}
# font_import() # first time on a machine only.  Watch the "Console" window for a confirmation dialog

# The following lines create a table that has the number of breweries
# per state, then tidys up the columns into a format the usmaps library
# understands
brewery_count_by_state <- breweries %>% group_by(State) %>% summarise(count=n())

brewery_count_by_state$State = as.character(brewery_count_by_state$State)    # convert factors to characters
brewery_count_by_state$State = str_trim(brewery_count_by_state$State)        # remove leading space from string
brewery_count_by_state = brewery_count_by_state %>% mutate(state=State)      # Capitolize State
brewery_count_by_state = brewery_count_by_state[,3:2]                        # remove the old, non-capitolized state

color_boxes <- data.frame(y1=seq(0,40,10),                                   # this sets up an array of x and y coords, and colors to
                           y2=seq(10,50,10),                                 # zebra stripe the background later
                           color=rep(c("#BFEFFF20","#BFEFFF40"),
                                     length.out = 5))

# turn the states back into factors
brewery_count_by_state$state <- factor(brewery_count_by_state$state, levels = brewery_count_by_state$state[order(brewery_count_by_state$count)])

# read in the photo background
image <- readJPEG("beer_glass_background.jpg")
# scale the background for this chart
bg <- rasterGrob(image, height = unit(1.9, "npc"), x=0.5, y=0.2)

brewery_count_by_state %>%                                                 # send in the brewery data
  mutate(fullname = abbr2state(state)) %>%                                 # change the state abbr to full name
  ggplot() +                     # send the data in ascending order
  annotation_custom(bg, -Inf, Inf, -Inf, Inf) +                            # add the background image
  geom_bar(aes(x=reorder(fullname, count), y = count), 
           stat="identity", fill = "brown", color = "gold", alpha = 0.9) + # plot the actual bars
  geom_text(aes(x=reorder(fullname, count), y = count,label = count), 
            vjust = -0.5, size = 3) +                                      # draw the count up above the bars
  geom_rect(data=color_boxes, aes(ymin = y1, ymax = y2, 
                    xmin = -Inf, xmax = +Inf), fill = color_boxes$color) + # zebra stripe the background
  theme_minimal() +                                                        # clear out the theme
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_text(family = "Georgia", size = 10),              # set text attributes 
        axis.text.x = element_text(family = "Verdana", angle=90, hjust = 1, vjust = .4)) +
  labs(title = "Breweries per State", x="", y="Number of Breweries")            # set legend
```

To perform the rest of the analysis, the data was checked for a number of inconsistencies including missing values.  Over half of the records(1405 of 2410) in the dataset did not have values for the IBU, and 62 of the entries did not have a value for ABV.

To allow us to continue without removing any of the records for missing values, it was decided to impute the data with a k-NN on beers in the same "Style".  What this means is that for a given beer, the algorithm will look at the collection of beers, find ones that match with it best, determine an appropriate IBU or ABV value, and replace the missing value in the sample.


```{r impute data}
# use the knnImputation() function from the Date Mining with R (DMwR) library to fill in missing data
# this method uses the available information in the remainder of the data frame to establish a kNN 
# relationship and fill in missing values
adjData <- data %>% knnImputation()
```

```{r median ABV per state}
# read in the photo background
image <- readJPEG("beer_glass_background.jpg")
# scale the background for this chart
bg <- rasterGrob(image, height = unit(1.5, "npc"), x=0.5, y=0.5)

# create the alternating color background
color_boxes <- data.frame(y1=seq(0.00, .06, .02), 
                          y2=seq(0.02, .08, .02),
                          color=rep(c("#BFEFFF20","#BFEFFF40"),
                                   length.out = 4))

adjData %>% 
  group_by(State) %>% 
  summarize(medabv=median(ABV)) %>%                                     # create a column named medabv from the mediab abv
  mutate(State = abbr2state(str_trim(as.character(State)))) %>%         # turn 2 letter factor into full state name
  ggplot() +
    annotation_custom(bg, -Inf, Inf, -Inf, Inf) +                       # beer background
    geom_bar(aes(x=reorder(State, medabv), y=medabv),                   # plot the actual data
             stat="identity", color="gold", fill="brown") +
    geom_rect(data=color_boxes,                                         # zebra stripe the background
           aes(ymin = y1, ymax = y2, xmin = -Inf, xmax = +Inf),
           fill = color_boxes$color) +
  scale_y_continuous(breaks = sort(c(seq(0.035, 0.065, length.out = 3 ), # setup the y axis scale, and add 0.62
                                     0.062, 0.04))) +                   # and 0.04 in it for reference
  geom_hline(yintercept = 0.062) +                                      # highlight the max at 0.062 abv
  geom_hline(yintercept = 0.050) +                                      # highlight the med at 0.050 abv
  geom_hline(yintercept = 0.04) +                                       # highlight the min at 0.04 abv
    theme_minimal() +                                                   # clear the theme
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_text(family = "Georgia", size = 10),      # set text attributes
        axis.text.x = element_text(family = "Verdana", angle=90, vjust = .4, hjust = 1)) + # rotate the x axis labels 90 degress
    labs(title = "State Median Alcohol By Volume", x="", y="Alcohol By Volume")  +         # set the graph labels
    coord_cartesian(ylim = c(0.035, 0.065))
```

```{r median IBU per state}

# create the alternating color background
color_boxes <- data.frame(y1=seq(0, 55, 10), 
                          y2=seq(10, 65, 10),
                          color=rep(c("#BFEFFF20","#BFEFFF40"),
                                   length.out = 6))


adjData %>% group_by(State) %>% 
  summarize(Median=median(IBU)) %>% 
  mutate(State = abbr2state(str_trim(as.character(State)))) %>%         # turn 2 letter factor into full state name
  ggplot() +
    annotation_custom(bg, -Inf, Inf, -Inf, Inf) +                       # beer background
    geom_bar(aes(x=reorder(State, Median), y=Median),
      stat="identity", color="gold", fill="brown") +
    geom_text(aes(x=reorder(State, Median), y = Median,label = round(Median), angle = 90), 
      hjust = -.5, size = 3) + 
    geom_rect(data=color_boxes,                                         # zebra stripe the background
           aes(ymin = y1, ymax = y2, xmin = -Inf, xmax = +Inf),
           fill = color_boxes$color) +
    theme_minimal() +                                                   # clear the theme
    theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_text(family = "Georgia", size = 10),      # set text attributes
        axis.text.x = element_text(family = "Verdana", angle=90, vjust = .4, hjust = 1)) + # rotate the x axis labels 90 degress
    labs(title = "State Median IBU", x="", y="IBU") +
    coord_cartesian(ylim = c(0, 65))
```

```{r highest IBU per state}

# create the alternating color background
color_boxes <- data.frame(y1=seq(0, 150, 50), 
                          y2=seq(50, 200, 50),
                          color=rep(c("#BFEFFF20","#BFEFFF40"),
                                   length.out = 4))


adjData %>% group_by(State) %>%
  summarize(Max=max(IBU)) %>% 
  mutate(State = abbr2state(str_trim(as.character(State)))) %>%         # turn 2 letter factor into full state name
  ggplot() +
    annotation_custom(bg, -Inf, Inf, -Inf, Inf) +                       # beer background
    geom_bar(aes(x=reorder(State, Max), y=Max),
      stat="identity", color="gold", fill="brown") +
    geom_rect(data=color_boxes,                                         # zebra stripe the background
           aes(ymin = y1, ymax = y2, xmin = -Inf, xmax = +Inf),
           fill = color_boxes$color) +
    geom_text(aes(x=reorder(State, Max), y = Max,label = round(Max), angle = 90), 
      hjust = -.5, size = 3) +
    theme_minimal() +                                                    # clear the theme
    theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_text(family = "Georgia", size = 10),                        # set text attributes
        axis.text.x = element_text(family = "Verdana", angle=90, vjust = .4, hjust = 1)) +# rotate the x axis labels 90 degress
  labs(title = "State Max IBU", x="", y="IBU")

```
```{r highest ABV per State }
# create the alternating color background
color_boxes <- data.frame(y1=seq(0.00, .12, .02), 
                          y2=seq(0.02, .14, .02),
                          color=rep(c("#BFEFFF20","#BFEFFF40"),
                                   length.out = 7))

adjData %>% 
  group_by(State) %>% 
  summarize(Max=max(ABV)) %>%
  mutate(State = abbr2state(str_trim(as.character(State)))) %>%         # turn 2 letter factor into full state name
  ggplot() +
  annotation_custom(bg, -Inf, Inf, -Inf, Inf) +                       # beer background
  geom_bar(aes(x=reorder(State, Max), y=Max),
      stat="identity", color="gold", fill="brown") +
  scale_y_continuous(breaks = sort(c(0.00, 0.06, 0.10, 0.14))) +               # setup the y axis scale, and add 0.62
                                                                       # and 0.04 in it for reference
  geom_hline(yintercept = 0.14) +                                      # highlight the max at 0.062 abv
  geom_hline(yintercept = 0.10) +                                      # highlight the med at 0.050 abv    
  geom_hline(yintercept = 0.06) +                                      # highlight the min at 0.04 abv
                                                                       
  geom_rect(data=color_boxes,                                          # zebra stripe the background
      aes(ymin = y1, ymax = y2, xmin = -Inf, xmax = +Inf),
      fill = color_boxes$color) + labs(title="Max ABV per State", y="ABV", x="State") +
  theme_minimal() +                                                    # clear the theme
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_text(family = "Georgia", size = 10),                        # set text attributes
        axis.text.x = element_text(family = "Verdana", angle=90, vjust = .4, hjust = 1))  # rotate the x axis labels 90 degress
```

The summary statistics of the ABV by state show a minimum ABV of 0.001, or 0.1% Alcohol by Volume.  This is either a data collection error or a set of Non-Alcoholic beers. The maximum ABV is 12.8%, and the mean ABV is 5.9% with 50% of the beers in the sample ranging from 5.0% to 6.7%.  A case could be made to discard the beers with minimum ABV as "not-beers" but rather "beer-like-beverages", as they are not subject to the same laws regulating the sale of alcoholic beverages.  Alternately, many persons are not allowed to drink alcohol because of religious beliefs or medical reasons, so it may be worth while to keep the non-alcoholic beers in the study.
```{r summary stats of abv}
adjData %>% select(ABV) %>% summary()
```
The distribution of Alcohol by Volume peaks at 5.9%, and shows a right skew, indicating that the set of beers sampled show a tendency to start at around 5% ABV, with a large grouping of beers in that 5.0%-6.7%, then a long tail reaching into higher and higher ABV numbers.  Fewer beers were below the 5.0% ABV mark.  This shows a trend towards making beers with higher and higher ABV, indicating a potential untapped market at lower alcohol contents.
```{r histogram of ABV, large binwidth}
adjData %>% filter(!is.na(ABV)) %>% 
  ggplot(aes(x=ABV)) +
  annotation_custom(bg, -Inf, Inf, -Inf, Inf) +                       # beer background
  geom_histogram(aes(y=..density..),binwidth = 0.005, na.rm = TRUE, show.legend = TRUE) +
  labs(title="Frequency of ABV", y="ABV", x="State") +
  geom_density(alpha=0.2, fill="#FF6666", color="red")
```


Using the original data set to plot the ABV vs. IBU with a linear model regression line, we see an "up and to the right" correlation between IBU and ABV.  Pearson's Correlation is 0.67, indication a fair amount of positive correlation.
```{r relationship between IBU and ABV}
data %>% select(c(IBU, ABV)) %>% ggplot() +
  annotation_custom(bg, -Inf, Inf, -Inf, Inf) +                       # beer background
  geom_point(aes(x=IBU, y=ABV), position = "jitter", na.rm = TRUE) +
  geom_smooth(aes(x=IBU, y=ABV), method="lm", na.rm = TRUE) +
  stat_cor(aes(x=IBU, y=ABV),method = "pearson", label.x = 100, label.y = 0.010, na.rm = TRUE) +
  annotate("text", x=117, y=0.017, label="Pearson's Correlation") +
  labs(title = "Correlation of ABV vs. IBU", x="IBU", y="ABV")
```
If we use the imputed dataset however, the additional data tends toward the mean and the correlation drops to 54%.  Even though it lost 13% correlation points, it is still a significant positive correlation.
```{r relationship between adjusted IBU and ABV}
adjData %>% select(c(IBU, ABV)) %>% ggplot() +
  annotation_custom(bg, -Inf, Inf, -Inf, Inf) +                       # beer background
  geom_point(aes(x=IBU, y=ABV), position = "jitter", na.rm = TRUE) +
  geom_smooth(aes(x=IBU, y=ABV), method="lm", na.rm = TRUE) +
  stat_cor(aes(x=IBU, y=ABV),method = "pearson", label.x = 100, label.y = 0.010, na.rm = TRUE) +
  annotate("text", x=117, y=0.017, label="Pearson's Correlation") +
  labs(title = "Correlation of ABV vs. IBU", x="IBU", y="ABV")
```


```{r Unnamed data exploration, include=FALSE}
adjData %>% 
  group_by(Brewery) %>% 
  summarize(count=n()) %>% 
  filter(count > 19) %>%  
  ggplot() + 
    geom_bar(aes(x=reorder(Brewery, count), y=count),stat = "identity") + 
    theme(axis.text.x = element_text(family = "Verdana", angle=90))
```

```{r knn classification for IBU of Ales}
#install.packages("ggvis")
library(ggvis)
library(class)
library(caret)

# we will use the un-imputed data for this, as the data is much more accurate
# pull out the IPAs first
ipas <- data %>% 
  filter(!is.na(IBU) & !is.na(ABV)) %>% 
  filter(str_detect(Style, regex("\\bipa\\b", ignore_case = TRUE, multiline = FALSE))) %>% 
  mutate(ipa=TRUE)

# then the "other" ales
otherAles <- data %>% 
  filter(!is.na(IBU) & !is.na(ABV)) %>% 
  filter(str_detect(Style, regex("\\bale\\b", ignore_case = TRUE, multiline = FALSE))) %>% 
  filter(!str_detect(Style, regex("\\bipa\\b", ignore_case = TRUE, multiline = FALSE))) %>% 
  mutate(ipa=FALSE)

allAles <- rbind(ipas, otherAles)

```
```{r draw ellipse around centroid}
if(!require(ggforce)) {install.packages("ggforce")}
if(!require(concaveman)) {install.packages("concaveman")}
library(ggforce)
library(concaveman)

allAles %>% ggplot(aes(x=IBU, y=ABV)) +
  annotation_custom(bg, -Inf, Inf, -Inf, Inf) +                       # beer background
  geom_point(aes(color = ipa)) +
  geom_density2d(aes(color=ipa)) +

  labs(title = "IBU x ABV for IPAs and Other Ales", color = "IPA?")

```
If we map the IBU vs. the ABV for the groups "IPA" and "Others", then add contour lines to show the greatest concentration of values, we can clearly see the groupings of IBU and ABV for Others to be much lower in IBU and ABV than the IPAs.  While we can see that there are clearly some Other ales in the region associated with IPAs, we'd like to know if the two groups are distinct enough that we can predict the Style of beer from the IBU/ABV alone.

```{r summary statistics on IBU and ABV}

summary(ipas$IBU)
summary(otherAles$IBU)
summary(ipas$ABV)
summary(otherAles$ABV)

ipas %>% ggplot(aes(x=IBU, y=ABV)) +
  geom_point() + geom_smooth(method="glm") + 
  stat_cor(aes(x=IBU, y=ABV),method = "pearson", label.x = 100, label.y = 0.010, na.rm = TRUE) +
  labs(title = "India Pale Ale IBU/ABV Correlation")

otherAles %>% ggplot(aes(x=IBU, y=ABV)) + 
  geom_point() + geom_smooth(method="glm") + 
  stat_cor(aes(x=IBU, y=ABV),method = "pearson", label.x = 80, label.y = 0.010, na.rm = TRUE) +
  labs(title = "Non-IPA IBU/ABV Correlation")
```


```{r run a 90% train/test knn on ipas}
# run a set of tests to determine best k value
iterations = 100 # the number of tests to run
k_runs = 50 # the number of "k"s to run in each test

# a place to hold the master data
masterAcc = matrix(nrow = iterations, ncol=k_runs)

for(j in 1:iterations) {
  # accs = data.frame(accuracy = numeric(k_runs), k=numeric(k_runs))

  # randomize the sample set for this test
  ran <- sample(1:nrow(allAles), 0.9 * nrow(allAles))
  ipa_train <- allAles[ran,]
  ipa_test <- allAles[-ran,]
  
  # run the 'k's
  for(i in 1:k_runs) {
    classifications = knn(ipa_train[,c(4,5)], ipa_test[,c(4,5)], ipa_train$ipa, k=i, prob = TRUE)
    CM = confusionMatrix(table(ipa_test$ipa, classifications))
    masterAcc[j,i] = CM$overall[1]
  }
}
```
```{r plot the mean accuracies from the knn test}

MeanAcc = colMeans(masterAcc)
plot(seq(1,k_runs,1),MeanAcc, type="l")




# run a single test with the experimentally gathered 'k' value
ipa_pr <- knn(ipa_train[,c(4,5)], ipa_test[,c(4,5)], ipa_train$ipa, k=5, prob = TRUE)

# get the statistics of the 
confusionMatrix(table(ipa_test$ipa, ipa_pr))



```
```{r internal cv knn}
set.seed(1972)
iterations = 100
k_runs = 50

masterAcc = matrix(nrow = iterations, ncol = k_runs)
for(j in 1:iterations) {
  accs = data.frame(accuracy = numeric(k_runs), k = numeric(k_runs))
  for(i in 1:k_runs) {
    classifications = knn.cv(allAles[,c(4,5)], allAles$ipa, k=i, prob = TRUE)
    CM = confusionMatrix(table(allAles$ipa, classifications))
    masterAcc[j,i] = CM$overall[1]
  }
}

MeanAcc = colMeans(masterAcc)
plot(seq(1,k_runs,1), MeanAcc, type="l")

```

```{r take the k from knn.cv and get actual parameters}
set.seed(1972)
classifications = knn.cv(allAles[,c(4,5)], allAles$ipa, k=5, prob = TRUE)
confusionMatrix(table(allAles$ipa, classifications))
```

The Internal Cross Validation k-Nearest Neighbor classification tells us that there is enough difference in the IBU/ABV of India Pale Ales (IPAs) to determine the Style of beer roughly 86% of the time, just from the values of IBU and ABV.  We have confidence that 95% of the time, the average accuracy of the kNN model used would be between 84% and 88% accuracy.


```{r investigate home state from IBU, ABV, Style}

# we must recode string factors as numbers for the distance algorithm in
# the kNN classifier to work properly
allAles <- allAles %>% 
  mutate(city_id = as.integer(City)) %>% 
  mutate(style_id = as.integer(Style))

classifications = knn.cv(allAles[,c(4,5,13)], allAles$State, k=1, prob = TRUE)
CM = confusionMatrix(table(allAles$State, classifications))
CM$overall[1]
```
If we assume that:

* current breweries are selling most of their beer locally
* that they are making beers which sell well in their areas

It could be beneficial to know what types of beer would sell best in a given state based on what is currently being produced in that state. Using a kNN classification that takes into account IBU, ABV and Style of beer, we can predict with around a 30% success rate the state that beer would most fit into.  30% may not sound like much, but compared to a purely random 1-in-50 chance of picking the right state (a 2% probability), 30% is a 15x improvement.

